The Rockerverse: Packages and Applications for Containerization with R

### 2020-01-28 manuscript.Rmd
reorder use cases, make R packages own chapter
also remove some unneeded fields in bib file
more explicit package imports, set locale to fix encoding warning in tex

```{r setup, include=FALSE}
library(tidyverse)
```

@@

# Use cases and applications
<!-- from packages closer to Docker to second level packages to more complex use cases -->
\label{applications}

@@

## Interfaces for Docker in R
\label{interfaces}

```{r clientfunctionalities, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
functionalities <- tribble(
  ~id, ~Functionality,
  "dockerfile", "Generate a Dockerfile",
  "build", "Build an image",
  "execute", "Execute a container locally or remotely",
  "cloud", "Deploy or manage an instances in the cloud",
  "interact", "Interact with an instance (e.g., file transfer)",
  "images", "Manage storage of images",
  "othercontainers", "Supports Docker and Singularity",
  "api", "Direct access to Docker API instead of using the CLI",
  "install", "Installing Docker software"
)

@@

Several second order R packages attempt to make the process of creating Docker images and using containers for specific tasks, such as running tests or rendering reproducible reports, easier.
While authoring and managing an environment with Docker by hand is possible and feasible for experts\footnote{See, e.g., this tutorial by RStudio on how to manage environments and package versions and to ensure deterministic image builds with Docker: \href{https://environments.rstudio.com/docker}{https://environments.rstudio.com/docker}.}, the following examples show the power of automation when environments become too cumbersome.
Especially _version pinning_, with packages \pkg{remotes} and \pkg{versions} for R or by using MRAN, and with system package managers for different operating systems, can greatly increase the reproducibility of built images and are common approaches.

@@

## Using R to power enterprise software in production environments
\label{enterprise}

@@

The prevelance of Docker in industry naturally leads to usage of containers with R in such settings as well, as customers already manage platforms in Docker containers.
These products often entail a high amount of open source software in combination with proprietary layers adding the relevant commercializable features.
One such example is RStudio's data science platform [RStudio Team](https://rstudio.com/products/team/).
It allows teams of data scientists and their respective IT/DevOps groups to develop and deploy code in R and Python around the RStudio Open Source Server inside of Docker images, without requiring users to learn new tools or directly interact with containers.
The best practices for [running RStudio with Docker containers](https://support.rstudio.com/hc/en-us/articles/360021594513-Running-RStudio-with-Docker-containers) as well as [Docker images](https://github.com/rstudio/rstudio-docker-products) for RStudio's commerical products are publicly available.

@@

## Processing
\label{processing}

@@

# Conclusions
\label{conclusions}

>>>>

```{r setup, include=FALSE}
Sys.setlocale('LC_ALL','C')
```

@@

# Interfaces for Docker in R
\label{interfaces}

```{r clientfunctionalities, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library("tibble")
library("tidyr")
library("dplyr")

functionalities <- tribble(
  ~id, ~Functionality,
  "dockerfile", "Generate a Dockerfile",
  "build", "Build an image",
  "execute", "Execute a container locally or remotely",
  "cloud", "Deploy or manage an instances in the cloud",
  "interact", "Interact with an instance (e.g., file transfer)",
  "images", "Manage storage of images",
  "othercontainers", "Supports Docker and Singularity",
  "api", "Direct access to Docker API instead of using the CLI",
  "install", "Installing Docker software"
)

# Use cases and applications
\label{applications}

@@

@@

Community maintained images provide a solid basis for extension as part of individual requirements.
Several second order R packages attempt to make the process of creating Docker images and using containers for specific tasks, such as running tests or rendering reproducible reports, easier.
While authoring and managing an environment with Docker by hand is possible and feasible for experts\footnote{See, e.g., this tutorial by RStudio on how to manage environments and package versions and to ensure deterministic image builds with Docker: \href{https://environments.rstudio.com/docker}{https://environments.rstudio.com/docker}.}, the following examples show the power of automation when environments become too cumbersome.
Especially _version pinning_, with packages \pkg{remotes} and \pkg{versions} for R or by using MRAN, and with system package managers for different operating systems, can greatly increase the reproducibility of built images and are common approaches.

@@

## Processing

@@

The prevelance of Docker in industry naturally leads to usage of containers with R in such settings as well, as customers already manage platforms in Docker containers.
These products often entail a high amount of open source software in combination with proprietary layers adding the relevant commercializable features.
One such example is RStudio's data science platform [RStudio Team](https://rstudio.com/products/team/).
It allows teams of data scientists and their respective IT/DevOps groups to develop and deploy code in R and Python around the RStudio Open Source Server inside of Docker images, without requiring users to learn new tools or directly interact with containers.
The best practices for [running RStudio with Docker containers](https://support.rstudio.com/hc/en-us/articles/360021594513-Running-RStudio-with-Docker-containers) as well as [Docker images](https://github.com/rstudio/rstudio-docker-products) for RStudio's commerical products are publicly available.
## Using R to power enterprise software in production environments
\label{enterprise}

@@

@@

# Conclusions
